Country-Level Impact of Education on Entrepreneurship

### 2014-12-12 PresentationandPaper/FromEducationToEntrepreneurship.Rmd
Final commit

@@

The support of students' self-confidence within education systems as we just defined it can also affect the students' entrepreneurial self-efficacy, that should be controlled for. To do so, we will look at the average perceived entrepreneurial capabilities and fear of failure in starting a business across the population at the country-level.

@@

Regarding potential control variables, we should normally consider including at least the main environmental determinants of entrepreneurship but the lack of consistency in the data at hand for our explanatory variables made us use a summary of those as a proxy. We chose to use the level of good opportunities to start a business perceived on average across the population of the given country.

@@

*Data Sources*

@@

<br>

@@

Additional information on the GEM data used for this research can be seen in the descriptive statistics table included below.

@@

@@

image: ![Way of Teaching](https://raw.githubusercontent.com/SamDund/Assign2Education/master/PresentationandPaper/images/WoTmap.png)

@@

-------------------------

@@

Given our data characteristics and the available models in the Zelig package, we used a "Least Squares Regression for Continuous Dependent Variables." A regression output table and our simulation results are included below.

@@

As we can see from the regression output table, using the Amelia program to generate additional data expanded the observations analyzed from 172 to 318, and increased the statistical significance of our explanatory variable.  In addition, the positive slope of this equation is consistent with what we expected after our analysis of existing literature on this topic.  In addition, this process reduced the variance in our simulation results, suggesting it improved the accuracy of this model.  Despite this, we have several concerns regarding the reliability of existing GEM data (which will be discussed later in this paper), which makes us question the effectiveness of using this data to generate additional observations.

Perceived opportunities

@@

**Results**

The GEM National Expert Survey however interesting and innovative for its extensive look into new sorts of data still allows for much improvement. It is essentially supposed to help control for the environmental determinants of entrepreneurship, which are here called the framework conditions. The major issue is not so much what it aims at revealing of the entrepreneurial context of each country but it has much more to do with methodology of data collection. Indeed, for each of these framework conditions, only four experts answered the survey, which makes the measure be very dependent on the individual respondent. This leads to highly variant data. Besides, the national teams hired by GEM Coordination Team were reported by GEM itself to manage time allocated for data collection rather poorly. That is, while they had five months at their disposal to approach and survey four experts for each framework condition, they happened on a regular basis to be in a hurry in July and therefore ended up relying on their personal connections to answer the surveys.

Regarding the D01 indicator that we used as a proxy for a way of teaching supportive of students' self-confidence, it is quite variant from year to year, which is not quite what we would expect. Indeed, as we look at the way of teaching and education features, these are generally culture related and as such should be much less fluctuant. At most, they should vary due to reforms being implemented at national, federal or state level, but that does not take place every year, rather every four or five years, sometimes only once a decade or even less frequently. Thus this yearly change is very unlikely to be a faithful figuration of how supportive of students' self-confidence teaching is like in a given country. On top of that, we purposely selected one item among the six used to characterised potential entrepreneurial education. That one item cannot be compensated by others in our study, which probably worsened the effect. The scale between 0 and 5 makes the assessment even more approximate. As a result of all these flaws combined leads to some variations of up to 1 point for some countries, like for Denmark for instance.

Now looking at GEM Adult Population Survey (APS), even if the data is directly collected by GEM from representatives samples of the population (around 2000 peole), once again we observe huge variation for indicators such as the perceived opportunities, which are unlikely to fluctuate that much in reality. That very indicator was used as a proxy for many potential control variables regarding economic outlook and other environmental determinants of entrepreneurship, and we do not expect the level of financing, the level of public programmes or the GDP growth for instance to vary that much from year to year. The subjective perception this indicator entails should have been compensated by the representativity of the sample and thus should not show so much variance. Once more, even if we are very doubtful about that data.

Last but not least is the scarcity of the data. For the perceived capabilities, perceived opportunities and fear of failure, there are already a few missing data but this gets much worse regarding the way of teaching. For that key explanatory variable, many countries have only data for have the years we were looking into, those who had the most data were at least missing three years, and three countries (Estonia, Luxembourg and Slovak Republic) had no data available at all.

Adding up all these limitations of the data we used, it does not come as a surprise that we could hardly get significant coefficients and could only observe general trends.

@@

Education systems tend to be assessed in terms of numbers of students, ratios of teachers among other quantitative measures but what we tried to suggest here is that they could be looked at in quite a different way. As governments allocate public funds for education, they happen to make a long term investment, and, although this may not come as news, we intended here to propose a different rationale in doing so. Without denying the importance of all the classic determinants of entrepreneurship, we suggest that education systems could turn out to be one of those determinants that favour start up activites.

From the research we attempted to conduct here, due to poor quality and inconsistent data, we cannot ascertain our point but we still firmly believe there could be something happening out there and this should be looked into more. Now that would mean collecting the adequate amount of data and computing a new kind of indicators that are not yet available. This surely is a prospect for future research.

>>>>

The process of allocating public funds for education is an opportunity to make calculated long-term investments in a nation's populace. The rationale for decisions made during this process, however, is often difficult to understand. That is, while many claims have been made in entrepreneurship literature regarding the importance of cultural and educational factors - not just structural and economic factors - in determining individual and societal tendencies to start a business, these factors are often not included in the decision-making process. Actually, only a few empirical studies have been carried out on this topic and most of them focused on which attitudes were favourable to starting a business, while almost none explored what determined these dispositions.

We believe that education systems can favour entrepreneurship and, beyond, GDP growth at the country-level, depending on how supportive of students' self-confidence and how socially inclusive they are. We see this process as many-fold but we only focus here on one of the potential causal relations. We want to understand specifically to want extent the support of individual self-confidence within education systems can affect the likelihood to become an entrepreneur, and thus impact the rate of entrepreneurship at the country level.

We use the Global Entrepreneurship Monitor data, both the Adult Population Survey (APS) and the National Expert Survey (NES) to characterise the features of education systems across OECD countries from 2001 until 2010. We control for the average entrepreneurial self-efficacy across the population as well as for the perceived opportunities to start a business. Due to scarce data, due to some flaws in the collection, we cannot ascertain here that the causal chain we look into actually takes place but we firmly believe that is a need for another type of indicators to deepen the research in that direction.

\end{singlespace}

@@

The support of students' self-confidence within education systems as we just defined it can also affect the students' entrepreneurial self-efficacy, that should be controlled for. To do so, we will look at the average perceived entrepreneurial capabilities and "Fear of Failure" in starting a business across the population at the country-level.

@@

Regarding potential control variables, we should normally consider including at least the main environmental determinants of entrepreneurship, but the lack of consistency in the data at hand for our explanatory variables made us use a summary of those as a proxy. We chose to use the level of good opportunities to start a business perceived on average across the population of the given country.

@@

**Data Sources**

@@

@@

Additional information on the GEM data used for this research can be seen in the descriptive statistics table  below.

@@

# This map will not display in a static Latex document.  Instead, we have included a screanshot of the map as it appears in an html file.

@@

![Way of Teaching](https://raw.githubusercontent.com/SamDund/Assign2Education/master/PresentationandPaper/images/WoTmap.png)

@@

@@

Given our data characteristics and the available models in the Zelig package, we used a "Least Squares Regression for Continuous Dependent Variables." A regression output table and our simulation results are below.

@@

As we can see from the regression output table, using the Amelia program to generate additional data expanded the observations analyzed from 172 to 318, and increased the statistical significance of our explanatory variable.  In addition, the positive slope of this equation is consistent with what we expected after our analysis of existing literature on this topic.  In addition, this process reduced the variance in our simulation results, suggesting it improved the accuracy of this model.  Despite this, we have several concerns regarding the reliability of existing GEM data (which will be discussed in the next section), which makes us question the effectiveness of using this data to generate additional observations.

The GEM National Expert Survey however interesting and innovative for its extensive look into new sorts of data still allows for much improvement. It is essentially supposed to help control for the environmental determinants of entrepreneurship, which are here called the framework conditions. The major issue is not so much what it aims at revealing of the entrepreneurial context of each country but it has much more to do with methodology of data collection. Indeed, for each of these framework conditions, only four experts answered the survey, which makes the measure be very dependent on the individual respondent. This leads to highly variant data. Besides, the national teams hired by GEM Coordination Team were reported by GEM itself to manage time allocated for data collection rather poorly. That is, while they had five months at their disposal to approach and survey four experts for each framework condition, they happened on a regular basis to be in a hurry in July and therefore ended up relying on their personal connections to answer the surveys [@OECDProject].

Regarding the "Way of Teaching" variable that we used as a proxy for a way of teaching supportive of students' self-confidence, it is quite variant from year to year, which is not quite what we would expect. Indeed, as we look at the way of teaching and education features, these are generally culture related and as such should be much less fluctuant. At most, they should vary due to reforms being implemented at national, federal or state level, but that does not take place every year, rather every four or five years or even less frequently. Thus this yearly change is very unlikely to be an accurate indicator of how supportive of students' self-confidence teaching is like in a given country. On top of that, we purposely selected one item among the six used to characterise potential entrepreneurial education (Way of Teaching). Isolating that one item from the other 5 factors may have lessened the indicator's reliability. In addition, the scale between 0 and 5 makes the assessment even more approximate. As a result of all these flaws combined leads to some variations of up to 1 point for some countries, including Denmark.

Now looking at GEM Adult Population Survey (APS), even if the data is directly collected by GEM from representatives samples of the population (around 2000 peole), once again we observe huge variation for indicators such as the "Perceived Opportunities," which are unlikely to fluctuate that much in reality. That very indicator was used as a proxy for many potential control variables regarding economic outlook and other environmental determinants of entrepreneurship, and we do not expect the level of financing, the level of public programmes or the GDP growth for instance to vary that much from year to year. The subjective perception this indicator entails should have been balanced by the representativity of the sample and thus should not show so much variance. Once more, even if we are very doubtful about that data.

A final issue has been the scarcity of the data. For the "Perceived Capability", "Perceived Oportunities" and "Fear of Failure" indicators, there are already several missing observations.  However, for our explanatory variable, "Way of Teaching," there is considerably more missing obersvations.  In fact, many countries have only data a few of the years we were looking into.  Those who had the most data were at least missing three years, and three countries (Estonia, Luxembourg and Slovak Republic) had no data available at all.  We have included a chart of "Perceived Opportunties" below to illustrate how this data has been inconsistent over time.

@@

Considering these data limitations, it does not come as a surprise that we struggled to obtain significant parameters and could only observe general trends.

@@

Education systems tend to be assessed in terms of numbers of students, ratios of teachers among other quantitative measures, but what we tried to suggest here is that they could be looked at in quite a different way. As governments allocate public funds for education, they happen to make a long term investment, and, although this may not come as news, we intended here to propose a different rationale in doing so. Without denying the importance of all the classic determinants of entrepreneurship, we suggest that education systems could turn out to be one of those determinants that favour start up activites.

From the research we attempted to conduct here, due to poor quality and inconsistent data, we cannot ascertain our point, but still firmly believe there is an under recognized impact of education on entrepreneurship and this should be more thoroughly explored. This would require additional collection and analysis of data and the creation of new kinds of indicators that are not yet available. This surely is a prospect for future research.

**References**
